Competing Bandits:\\ The Perils of Exploration under Competition \thanks{This is an extended abstract of \citet{CompetingBandits-merged}, which in turn subsumes conference papers \citet{CompetingBandits-itcs18} and \citet{CompetingBandits-ec19}.}

### 2021-03-02 journal paper/content/related_work.tex
Minor edits in related work.

\asedit{The tension between exploration and incentives}
%The three-way tradeoff between exploration, exploitation and incentives 
has been studied in several scenarios very different from ours:
incentivized exploration in recommendation systems
(starting from \citet{Kremer-JPE14,Che-13}, see \citet[Ch. 11]{slivkins-MABbook}),
    %\citep[\eg][]{Che-13,Frazier-ec14,Kremer-JPE14,ICexploration-ec15,Bimpikis-exploration-ms17,Bahar-ec16,Jieming-unbiased18},
dynamic auctions
    \citep{DynAuctions-survey10},
    %\citep[\eg][]{AtheySegal-econometrica13,DynPivot-econometrica10,Kakade-pivot-or13},
pay-per-click ad auctions
    \citep[\eg][]{MechMAB-ec09,DevanurK09},
coordinating search and matching
    \citep{Bobby-Glen-ec16},
and human computation
    \citep[\eg][]{RepeatedPA-ec14,Ghosh-itcs13}.
%A literature review of this work can be found in
%\citep[Ch. 11.6]{slivkins-MABbook}.

\xhdr{Exploration and competition.}
\citet{bergemann1997market,bergemann2000experimentation,keller2003price} study the interplay of exploration and competition for users when the competing firms experiment with \emph{prices}
(whereas in our model they experiment with design alternatives).
All three papers consider strategies that respond to competition and analyze Markov-perfect equilibria (whereas we focus on the adoption of better bandit algorithms). We discuss them in more detail below.

@@

%\gaedit{\citet{bergemann1997market,bergemann2000experimentation} studied a model where there is an incumbent with known product quality and an entrant with unknown, but exogenously determined, quality. Similar to \citet{keller2003price}, the firms compete on prices and they characterize the MPE of dynamic competition between the two firms. In their model, consumer purchases of the entrant's product result in an informational externality that allows both the firms and consumers to gather information on the entrant's product quality. However, we focus on a model where firms' product quality is endogenous and the focus of exploration in a context where the price mechanism is absent.}
In the line of work on \emph{strategic experimentation} (starting from \citet{Bolton-econometrica99,Keller-econometrica05}, see \citet{Horner-survey16} for a survey), agents explore and learn over time in a shared environment. Thus, we have exploration algorithms which interact with each other strategically, \eg each agent prefers to free-ride on someone else's exploration. However, the agents do not compete with each other in any meaningful sense.

>>>>

\asedit{A monopolistic bandit algorithm may interact with self-interested parties, leading to a tension between exploration and incentives. This tension has been studied in several scenarios:}
%\asedit{The tension between exploration and incentives}
%The three-way tradeoff between exploration, exploitation and incentives
%has been studied in several scenarios very different from ours:
incentivized exploration in recommendation systems
(starting from \citet{Kremer-JPE14,Che-13}, see \citet[Ch. 11]{slivkins-MABbook}),
    %\citep[\eg][]{Che-13,Frazier-ec14,Kremer-JPE14,ICexploration-ec15,Bimpikis-exploration-ms17,Bahar-ec16,Jieming-unbiased18},
dynamic auctions
    \citep{DynAuctions-survey10},
    %\citep[\eg][]{AtheySegal-econometrica13,DynPivot-econometrica10,Kakade-pivot-or13},
pay-per-click ad auctions
    \citep[\eg][]{MechMAB-ec09,DevanurK09},
coordinating search and matching
    \citep{Bobby-Glen-ec16},
and human computation
    \citep[\eg][]{RepeatedPA-ec14,Ghosh-itcs13}.
\asedit{Unlike this work, we focus on incentives created in a competition.}
%A literature review of this work can be found in
%\citep[Ch. 11.6]{slivkins-MABbook}.

\xhdr{Exploration and competition.}
\asedit{Several papers consider exploration algorithms in scenarios when the explorer is not a monopolist. The technical models are very different, and not amenable to the high-level question articulated in the Introduction.}

Most relevant are \citet{bergemann1997market,bergemann2000experimentation,keller2003price}. They study the interplay of exploration and competition for users when the competing firms experiment with \emph{prices} (whereas in our model the firms experiment with design alternatives).
All three papers consider strategies that respond to competition and analyze Markov-perfect equilibria (whereas we focus on the adoption of better bandit algorithms). We discuss them in more detail below.

@@

%\gaedit{\citet{bergemann1997market,bergemann2000experimentation} studied a model where there is an incumbent with known product quality and an entrant with unknown, but exogenously determined, quality. Similar to \citet{keller2003price}, the firms compete on prices and they characterize the MPE of dynamic competition between the two firms. In their model, consumer purchases of the entrant's product result in an informational externality that allows both the firms and consumers to gather information on the entrant's product quality. However, we focus on a model where firms' product quality is endogenous and the focus of exploration in a context where the price mechanism is absent.}
In the line of work on \emph{strategic experimentation} (starting from \citet{Bolton-econometrica99,Keller-econometrica05}, see \citet{Horner-survey16} for a survey), agents explore and learn over time in a shared environment. Thus, we have exploration algorithms which interact with each other strategically, \eg each agent prefers to free-ride on someone else's exploration. \asedit{However, this work is all about cooperation (or lack thereof), rather than competition.}
%the agents do not compete with each other in any meaningful sense.
